---
title: "anonymize-data"
output: html_document
date: "2024-07-10"
---

```{r setup}

library(tidyverse)
library(readxl)
library(janitor)
library(digest)

```

# notes

For testing, in Micah's local environment we have options for list data:

- 'data_2024-12-14' which has all daily reports
- 'data_14_2024-01-14' which has 14 daily reports representing 2024-01-01 to 2024-01-14
    - When the first row has data (see [[1]]:[[3]]), dates are read in as dates, "double (S3: POSIXit, POSIXt)" 
    - When the first row is NA (see [[4]]:[[5]]), dates are in 5-digit format and read in as characters
- 'data_4_2024-01-13' which has 4 daily reports representing 2024-01-10 to 2024-01-14
    - All first rows are NA
- 'data_3_2024-01-03' which has 3 daily reports represention 2024-01-01 to 2024-01-03
    - All first rows have data

# import data

## error importing dates

NOTES
- with line ~105 included '`DATE OF BIRTH` = convert_to_date(`DATE OF BIRTH`),' error "not all strings converted to class date" with a list of 5-digit numbers
    - Stops run
- with line ~105 excluded '#`DATE OF BIRTH` = convert_to_date(`DATE OF BIRTH`),' error in bind_rows() can't combine date_of_birth '1$date_of_birth' (datetime<UTC> and '2$date_of_birth' <characer>) 
    - Stops run

POSSIBLE FIX: I think the fix is to add a line to the code cleaning reports with a line that converts excel 5-digit date time format to UTC datetime
    - https://rdrr.io/cran/janitor/man/excel_numeric_to_date.html 

PUZZLING: this is a problem for 'DATE OF BIRTH' specifically when all dates are the the same character or datetime format throughout any given monitoring report

````{r}

# create a list of df
# REPRODUCIBILITY: change local_path and file_names based on local environment
local_path <- "/Users/micahclarkmoody/Desktop/monitoring_data/data_2024-12-14/"
file_names <- list.files(path = local_path)
path_names <- paste(local_path, file_names, sep = "")

# import df in a list
list_data <- lapply(path_names, function(x) read_excel(paste0(path = x)))

# initialize an empty data frame for adding onto
final <- tibble()

```

```{r}

################################################################################
################################################################################
# alright, now we do the thing! loop over all files to clean & put them together
################################################################################
################################################################################

for (i in 1:length(list_data)) {
  
  print(file_names[i])
  print(i)

  ##############################################################################
  # set filename and date for loop
  ##############################################################################
  
  f <- list_data[[i]]
  date <- as.Date(substr(file_names[i], 1, 10))
  
  if (ncol(f) == 20) { # if there are 20 columns, then it's already in the format we want
    
    
    # since it's already in the format we want, we just need to clean names and 
    # add the date column which we already extracted from the file name
    f <- f %>%
      # clean column names
      clean_names() %>%
      # add snapshot date
      mutate(snapshot_date = date)
    
    # now that it's in the format we want, append to output dataframe
    final <- final %>% bind_rows(f)
    
  }
  
  # if it's not the above format, look for the next format:
  # if the first row has 27 NAs in it, then we know it's the format with 3 blank
  # rows at the top
  else if ( (list_data[[i]] %>% head(3) %>% mutate(countna = rowSums(is.na(.))) %>% pull(countna))[1] == 27 ) {
    
    f <- f %>%
      # get rid of empty columns
      select(-`...2`, -`...4`, -`...6`, -`...9`, 
             -`Court Compliance`, -`...18`, -`...24`) %>%
      # add col with sums of NAs, for filtering
      mutate(countna = rowSums(is.na(.))) %>%
      # get rid of the first 3 columns (which all have 20 NAs)
      filter(countna != 20)%>%
      # get rid of the countna column
      select(-countna) %>%
      # use janitor package to make first row rownames
      row_to_names(row_number = 1) %>%
      # use janitor to make date columns actually dates
      mutate(`ARREST DATE` = convert_to_date(as.numeric(`ARREST DATE`)),
             `BOOK DATE` = convert_to_date(as.numeric(`BOOK DATE`)),
             `DATE OF BIRTH` = convert_to_date(as.numeric(`DATE OF BIRTH`)),
             `RELEASE DATE` = convert_to_date(as.numeric(`RELEASE DATE`))) %>%
      # clean column names
      clean_names() %>%
      # booking_no column is character here, but numeric for the first filetype
      mutate(booking_no = as.numeric(booking_no),
             arrest_time = as.numeric(arrest_time),
             book_time = as.numeric(book_time),
             rlse_time = as.numeric(rlse_time)) %>%
      # add date column
      mutate(snapshot_date = date) 
    
    
    # now that it's in the format we want, append to output dataframe
    final <- final %>% bind_rows(f)
    
  }
  
  # if it's not in one of the two formats from above, then supposedly we know for
  # sure that it's the third format. If that's truly the case, we don't need
  # to test for it, and could just say else{}
  # 
  # I'm going to go ahead and be thorough and test for it
  # 
  # we know that column 1 is booking no & column 2 is "...2" so test for that:
  else if ( (names(list_data[[i]])[1] == "BOOKING NO." & names(list_data[[i]])[2] == "...2") ) {
    
    f <- f %>% 
      # remove columns where all rows are NA
      select(where(function(x) any(!is.na(x)))) %>%
      # clean column names
      clean_names() %>%
      # add date column:
      mutate(snapshot_date = date)
    
    
    final <- final %>% bind_rows(f)
    
  }
  
  # if none of these criteria are met, there's a problem, and we should know about it
  else {
    print(paste("oops! problem with ", file_names[i]))
  }
  
}

```

# Create hashed variables and delete identifiable variables

FLAG: This version of the code creates `id` based on FIRST NAME, LAST NAME, DATE OF BIRTH but not MIDDLE. This is because MIDDLE is missing for most observations. If, at any point, on wants to look at re-arrests including MIDDLE in the has would mess up that comparison because people with the same FIRST, LAST, DOB who were arrested once with MIDDLE recorded and once with MIDDLE=NA would appear as different people. To change this decision, add MIDDLE to line 'unite(id, c("last_name", "first_name"))'

```{r}

anon_final <- final %>%
  
  # create hash for name
  # flag, omitting MIDDLE
  unite(id, c("last_name", "first_name")) %>%
  rowwise() %>%
  mutate(id_hash = digest(id)) %>%
  ungroup() %>%
  
  # create hash for booking number
  rowwise() %>%
  mutate(booking_hash = digest(booking_no)) %>%
  ungroup() %>%
  
  # delete identifiable columns
  select(-c("booking_no", "id", "middle", "date_of_birth"))

```

# TEMPORARY test that anon_final is accurate



# Export data as a csv then email to Vera! 

```{r}

write.csv(anon_final, "/Users/micahclarkmoody/Desktop/anonymized_final.csv")

```






